---
title: "Laptop Price Prediction"
date: today
author: Tsung-Yu Lee
format:
 pdf:
    include-in-header:
      - text: |
         \usepackage{setspace,relsize}
mainfont: "Times New Roman"
toc: true
editor_options: 
  chunk_output_type: console
geometry:
  - top=10mm
  - right=10mm
  - left=10mm
  - bottom=10mm
---

\newpage

# Introduction

## Data Source

The laptop dataset is from Kaggle [Laptop Sales Price Prediction](https://www.kaggle.com/datasets/siddiquifaiznaeem/laptop-sales-price-prediction-dataset-2024).

```{r Read data, message=FALSE, warning=FALSE}
setwd("C:/github_LTY/homework/Final Report") # Set working directory.
laptop <- read.csv("data/laptop.csv", na.strings = c("","NA"))
```

Because "X" and "Name" are useless, so we delete these columns in advance.

```{r Delete X and Name}
laptop <- subset(x = laptop, select = -c(X, Name))
```

## Summary Statistics

```{r Summary Statistics, message=FALSE, warning=FALSE}
#| output: asis
library(Hmisc) # summary statistics
latex(describe(laptop), file = "", caption.placement = "top")
```

## Missing Values Proportion

```{r}
Summary.Table <- function(data){
  output <- data.frame(feature = names(data),
                       type = sapply(data, class),
            unique_value = sapply(data, function(x){if(!is.numeric(x)){length(unique(x))}
            else{"---"}}),
            num_missing = sapply(data, function(x){sum(is.na(x))}),
                       missing_rate = sapply(data, function(x) round(mean(is.na(x))*100, 2)))
  colnames(output) <- (c("feature", "type", "# of unique value",
                   "# of missing value", "missing rate(%)"))
  rownames(output) <- NULL
  return(output)
}
```

```{r Missing Values Proportion, message=FALSE, warning=FALSE}
library(DataExplorer)
library(gt)
Summary.Table(laptop) %>% gt()
```

## Missing Pattern

Next, we want to realize the missing pattern of laptop data.

```{r Define md.pattern function, message=FALSE, warning=FALSE}
# Customize md.pattern function originally in mice package.
library(mice)
md.pattern.mine <- function(x, plot = TRUE, rotate.names = FALSE, cex = 2.0){
  R <- is.na(x)
  # Calculate number of missing values in each variable.
  nmis <- colSums(R) 
  # create a matrix, and column is ordered by nmis.
  R <- matrix(R[, order(nmis)], dim(x))
  # Transform rows to strings consisting of 0(non-missing) and 1(missing).
  pat <- apply(R, 1, function(x) paste(as.numeric(x), collapse = ""))
  # Sort rows by their missing patterns.
  sortR <- matrix(R[order(pat), ], dim(x))
  # Keep non-duplicated rows.
  mpat <- sortR[!duplicated(sortR), ]
  # obtain number of each missing pattern at LHS.
  rownames(mpat) <- table(pat)
  # abs(mpat - 1) obtain entries of matrix, (0, 1) means (missing, non-missing)
  # rowSums(mpat) obtain number of missing values in each row at RHS.
  r <- cbind(abs(mpat - 1), rowSums(mpat))
  # obtain number of missing values in each column at bottom center.
  # sum(nmis) sums values at bottom center.
  r <- rbind(r, c(nmis[order(nmis)], sum(nmis)))
  if (plot){
    op <- par(mar = rep(0, 4))
    on.exit(par(op))
    plot.new()
    if (is.null(dim(R))){
      R <- t(as.matrix(R))
    }
    R <- r[1:nrow(r) - 1, 1:ncol(r) - 1]
    if(rotate.names){
      adj <- c(0, 0.5)
      srt <- 90
      length_of_longest_colname <- max(nchar(colnames(r)))/2.6
      plot.window(xlim = c(-1, ncol(R) + 1),
                ylim = c(-1, nrow(R) +
                        length_of_longest_colname), asp = 1)
    }
    else{
      adj <- c(0.5, 0)
      srt <- 0
      plot.window(xlim = c(-1, ncol(R) + 1), 
                  ylim = c(-1, nrow(R) + 1), asp = 1)
    }
    M <- cbind(c(row(R)), c(col(R))) - 1
    shade <- ifelse(R[nrow(R):1, ], mdc(1), mdc(2))
    rect(M[, 2], M[, 1], M[, 2] + 1, M[, 1] + 1, col = shade)
    for(i in 1:ncol(R)){
      # Show colnames at the top center.
      text(i - 0.5, nrow(R) + 0.3, colnames(r)[i], adj = adj, 
          srt = srt, cex = cex)
      # Show number of missing values for each variable at the bottom center.
      text(i - 0.5, -0.8, nmis[order(nmis)][i], cex = cex, srt = srt)
    }
    for(i in 1:nrow(R)){
      # Show number of missing values for each subjects at the RHS.
      text(ncol(R) + 0.3, i - 0.5, r[(nrow(r) - 1):1, ncol(r)][i], 
          adj = 0, cex = cex)
      # Show number of subjects for each kind of missing pattern at the LHS.
      text(-0.3, i - 0.5, rownames(r)[(nrow(r) - 1):1][i], adj = 1, cex = cex)
    }
    # show total number of missing entries.
    # text(ncol(R) + 1.3, -0.6, r[nrow(r), ncol(r)], cex = cex) 
    # return(r)
  }
  else {
    return(r)
  }
}
```

```{r, fig.height=7, fig.width=7, message=FALSE, warning=FALSE}
library(dplyr)
mis_col <- colnames(laptop)[profile_missing(laptop)$num_missing > 0]
laptop[,mis_col] %>%
  md.pattern.mine(rotate.names = T, cex = 1.0)
```

From the missing pattern plot, we can observe that  
1. The missing values of "Total_processor" and "Execution_units" occur with each other.  

# Exploratory Data Analysis (EDA)

## Numerical Columns

First, we look at numerical variables.
```{r}
numerical_columns <- laptop %>%
  select(where(is.numeric)) %>%
  colnames()
numerical_columns
```

Next, we look at the outcome variable ```Price```. Because the price in this data is INR (Indian Rupees), so we multiply 0.39, the exchange rate of TWD to INR.

```{r}
laptop$Price <- laptop$Price * 0.39
```

Because the original price is extremely skewed to the right, we take log() to make it more symmetric. Below are histograms of Price and log(Price). 
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
plot1 <- laptop %>%
  ggplot(aes(x = Price))+
  geom_histogram(binwidth = 10000, fill="lightblue",alpha=0.9)+
  ggtitle("Histogram of Price")

plot2 <- laptop %>%
  ggplot(aes(x = log(Price)))+
  geom_histogram(binwidth = 0.1, fill="lightblue",alpha=0.9)+
  ggtitle("Histogram of log(Price)")

grid.arrange(grobs = list(plot1, plot2), ncol = 2, nrow = 1)
```

Next, we look at the correlation between all numerical features.
```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
library(reshape2)
library(tidyverse)
# Compute the correlation matrix using pairwise deletion.
cor_matrix <- round(cor(laptop[,numerical_columns], use = "pairwise.complete.obs"), 3)

# Melt the correlation matrix for ggplot2
melted_cor_matrix <- melt(cor_matrix)

# Plotting with ggplot2
ggplot(data = melted_cor_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "black", linewidth = 0.5) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = value), color = "black", size = 6) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1)) +
  labs(title = "Correlation between numerical columns", x = "", y = "")
```

From the correlation heatmap, we can discover that  
1. ```"ppi"```, ```"Vertical_pixel"``` and ```"Horizontal_pixel"``` are highly correlated. We will keep ```"ppi"```, and delete ```"Vertical_pixel"``` and ```"Horizontal_pixel"```.  
2. ```"Energy_Efficient_Units"``` and ```"Low_Power_Cores"``` are perfectly correlated ($\rho=1$).  
We will keep ```"Energy_Efficient_Units"``` and delete ```"Low_Power_Cores"```. 
3. ```Rating``` has low correlation with ```Price```, so we will also delete it.


Also, we think that ```Processor_gen``` has wrong information about its meaning, we will also delete it.

```{r}
laptop <- subset(laptop, select = -c(Rating, Processor_gen, Low_Power_Cores,
                                     Energy_Efficient_Units, Vertical_pixel,
                                     Horizontal_pixel))
```

## Categorical Columns

We look at each categorical variables.  

```{r Categorical colomns}
# Obtain column names of categorical variables. 
cat_cols <- laptop %>%
  select(where(is.character)) %>%
  colnames()
cat_cols
```

1. Brand  
```{r Brand}
sort(table(laptop$Brand))
```

Because there is typing error about "ASUS" and "Asus", we unify them to "ASUS".  

```{r}
laptop[laptop == "Asus"] <- "ASUS"
```

2. Processor_brand  
```{r}
sort(table(laptop$Processor_brand))
```
3. Processor_name  
```{r}
sort(table(laptop$Processor_name))
```
Because there are so many names that have only 1 subject, we have to modify ```Processor_name``` by hand.  
```{r}
laptop[laptop == "AMD Athlon Pro"] <- "AMD Athlon"
laptop[laptop == "AMD Athlon Silver"] <- "AMD Athlon"
laptop[laptop == "eration Intel Core"] <- "Intel Core i5"
laptop[laptop == "Intel Celeron Dual"] <- "Intel Celeron"
laptop[laptop == "Intel Core 3"] <- "Intel Core i3"
laptop[laptop == "Intel Core I3-1115G4"] <- "Intel Core i3"  
laptop[laptop == "Intel Pentium "] <- "Intel Pentium"
laptop[laptop == "Intel Pentium Gold"] <- "Intel Pentium"
laptop[laptop == "Mediatek"] <- "MediaTek"
laptop[laptop == "Apple M2 Apple M2 Chip"] <- "Apple M2"
laptop[laptop == "Apple M3 Max"] <- "Apple M3"
laptop[laptop == "Intel  "] <- "Intel Celeron"
laptop[laptop == "Intel Pentium Silver"] <- "Intel Pentium" 
laptop[laptop == "Apple M3 Pro"] <- "Apple M3"
laptop[laptop == "Intel Core 7"] <- "Intel Core i7"
laptop[laptop == "MediaTek Kompanio"] <- "MediaTek"
laptop[laptop == "Intel Core 5"] <- "Intel Core i5"
laptop[laptop == "Intel Celeron "] <- "Intel Celeron"
```
4. Processor_variant  
```{r}
length(unique(laptop$Processor_variant))
```
Since ```Processor_variant``` has 126 categories, we don't need to explore this feature. We will delete this column in the data preprocessing stage.

5. RAM_type
```{r}
sort(table(laptop$RAM_type))
```
Because there are some typing errors, we modify them by hand.  
```{r}
laptop[laptop == "LPDDRX4"] <- "LPDDR4X"
laptop[laptop == "PDDR5X"] <- "LPDDR5X"
```
Also, there are 22 missing values in RAM_type, we surfed the internet, find the true RAM_type, and modified them by hand.  
```{r}
which(is.na(laptop$RAM_type))
laptop[c(18,84,117,164,203,204,295,348,349,350,351) , "RAM_type"] <- "Unified"
laptop[176, "RAM_type"] <- "LPDDR5"
laptop[296, "RAM_type"] <- "DDR4"
laptop[403, "RAM_type"] <- "DDR4"
laptop[552, "RAM_type"] <- "DDR4"
laptop[c(649,652), "RAM_type"] <- "Unknown"
laptop[657, "RAM_type"] <- "DDR4"
laptop[664, "RAM_type"] <- "DDR4"
laptop[728, "RAM_type"] <- "DDR4"
laptop[774, "RAM_type"] <- "DDR4"
laptop[903, "RAM_type"] <- "LPDDR5"
```
6. Storage_type  
```{r}
sort(table(laptop$Storage_type))
```

7. Graphics_name  
```{r}
length(unique(laptop$Graphics_name))
```
Since ```Graphics_name``` has 137 categories, we don't need to explore this feature. We will delete this column in the data preprocessing stage.

8. Graphics_brand
```{r}
sort(table(laptop$Graphics_brand))
```
Because there are 2 missing values in ```Graphics_brand```, we modified them by hand.
```{r}
which(is.na(laptop$Graphics_brand))
laptop[649, "Graphics_brand"] = "Unknown"
laptop[999, "Graphics_brand"] = "NVIDIA"
```
9. Graphics_integreted
```{r}
sort(table(laptop$Graphics_integreted))
```
Because there are 2 missing values in ```Graphics_integreted```, but we don't know the true value, we modified them by the majority class "False", then convert "True" to 1, "False" to 0.  
```{r}
which(is.na(laptop$Graphics_integreted))
laptop[c(649,999), "Graphics_integreted"] = "False"
laptop$Graphics_integreted <- ifelse(laptop$Graphics_integreted == "True", 1, 0)
```

10. Touch_screen
```{r Look at Touch_screen}
sort(table(laptop$Graphics_integreted))
```
We convert "True" to 1, "False" to 0.
```{r Convert Graphics_integrated to (0, 1)}
laptop$Touch_screen <- ifelse(laptop$Touch_screen == "True", 1, 0)
```
11. Operating_system
```{r}
sort(table(laptop$Operating_system))
```
Inside the Operating_system, there are some subjects that are too detailed, so we modified them to general ones.  
```{r}
laptop[laptop == "Android 11 OS"] <- "Prime OS"
laptop[laptop == "DOS 3.0 OS"] <- "DOS OS"
laptop[laptop == "Mac 10.15.3\t OS"] <- "Mac OS"
laptop[laptop == "Mac Catalina OS"] <- "Mac OS"
```
After we modify some typing errors, we delete ```Processor_variant```, ```Graphics_name``` for their high cardinalities, and delete ```Processor_brand``` since it's overlapped to another feature ```Processor_name```.
```{r}
laptop <- subset(laptop, select = -c(Processor_brand, Processor_variant, Graphics_name))
```

# Data Preprocessing
## Imputation
We use MICE imputation with method = 'pmm' to impute our missing data.    
```{r, message=FALSE, warning=FALSE}
temp_data <- mice(laptop, m=5, maxit=100, meth='pmm', seed=500, printFlag = FALSE)
data <- complete(temp_data, 1)
```

## Drop rare subjects
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
value_counts <- data %>% group_by(Brand) %>%  summarise(count = n())%>% arrange(count)
data <- data %>% group_by(Brand) %>% filter(n() >= 6) %>% ungroup()
value_counts <- data %>% group_by(Processor_name) %>%  summarise(count = n())%>% arrange(count)
data <- data %>% group_by(Processor_name) %>% filter(n() >= 2) %>% ungroup()
value_counts <- data %>% group_by(Graphics_brand) %>%  summarise(count = n())%>% arrange(count)
data <- data %>% group_by(Graphics_brand) %>% filter(n() >= 4) %>% ungroup()
value_counts <- data %>% group_by(RAM_type) %>%  summarise(count = n())%>% arrange(count)
data <- data %>% group_by(RAM_type) %>% filter(n() >= 2) %>% ungroup()
value_counts <- data %>% group_by(Storage_type) %>%  summarise(count = n())%>% arrange(count)
data <- data %>% group_by(Storage_type) %>% filter(n() >= 2) %>% ungroup()
value_counts <- data %>% group_by(Operating_system) %>%  summarise(count = n())%>% arrange(count)
data <- data %>% group_by(Operating_system) %>% filter(n() >= 4) %>% ungroup()
```
Finally, we look at the data we will analyze later.
```{r}
Summary.Table(data) %>% gt()
```

We split the dataset into training and testing data.
```{r}
sample_indices <- sample(seq_len(nrow(data)), size = 0.8 * nrow(data))
trainData <- data[sample_indices, ]
testData <- data[-sample_indices, ]
cat("Training set dimensions:", dim(trainData), "\n")
cat("Testing set dimensions:", dim(testData), "\n")
```

# Model Fitting
```{r}
MAE <- function(pred, true) return(mean(abs(pred-true)))
RMSE <- function(pred, true) return(sqrt(mean((pred-true)^2)))
```

## Linear Regression
```{r}
fit <- lm(log(Price) ~ ., data = trainData)
summary(fit)
```

```{r}
lr_pred <- exp(predict(fit, newdata = testData))
lr_mae <- MAE(lr_pred, testData$Price)
lr_rmse <- RMSE(lr_pred, testData$Price)
cat("LR MAE:", lr_mae, "RMSE", lr_rmse)
```

## Support Vector Regression
```{r, message=FALSE, warning=FALSE}
library(e1071)
svr_model <- svm(log(Price) ~ ., data = trainData, kernel = "linear")
svr_pred <- exp(predict(svr_model, newdata = testData))
svr_mae <- MAE(svr_pred, testData$Price)
svr_rmse <- RMSE(svr_pred, testData$Price)
cat("SVR MAE:", svr_mae, "RMSE", svr_rmse)
```

## Random Forest Regressor
```{r, message=FALSE, warning=FALSE}
library(randomForest)
rf_model <- randomForest(x = trainData[,-1], y = trainData$Price, importance = TRUE)
rf_pred <- predict(rf_model, newdata = testData[,-1])
rf_mae <- MAE(rf_pred, testData$Price)
rf_rmse <- RMSE(rf_pred, testData$Price)
cat("RF MAE:", rf_mae, "RMSE", rf_rmse)
```

## XGBoost
XGBoost and LightGBM need to onehot encoding for categorical data !  
```{r}
data_numerical <- data %>% select(where(is.numeric))
data_categorical <- data %>% select(where(is.character))
onehot.matrix <- model.matrix(~ 0 + Brand + Processor_name + RAM_type + Graphics_brand
                              + Operating_system, data = data_categorical)
data_ML <- cbind(data_numerical, onehot.matrix)
trainData <- data_ML[sample_indices, ]
testData <- data_ML[-sample_indices, ]
```

```{r, message=FALSE, warning=FALSE}
library(xgboost)
features_train <- trainData[, !colnames(trainData) %in% "Price"]
features_test <- testData[, !colnames(testData) %in% "Price"]
target_train <- trainData$Price
target_test <- testData$Price

dtrain <- xgb.DMatrix(data = as.matrix(features_train), label = log(target_train))
dtest <- xgb.DMatrix(data = as.matrix(features_test), label = log(target_test))
xgb_model <- xgboost(data = dtrain, nrounds = 100, objective = "reg:squarederror", verbose = 0)
xgb_pred <- exp(predict(xgb_model, dtest))
xgb_mae <- MAE(xgb_pred, testData$Price)
xgb_rmse <- RMSE(xgb_pred, testData$Price)
cat("XGB MAE:", xgb_mae, "RMSE", xgb_rmse)
```

## LightGBM
```{r, message=FALSE, warning=FALSE}
library(lightgbm)
features_train <- trainData[, !colnames(trainData) %in% "Price"]
features_test <- testData[, !colnames(testData) %in% "Price"]
target_train <- trainData$Price
target_test <- testData$Price

dtrain <- lgb.Dataset(data = as.matrix(features_train), label = log(target_train))
dtest <- lgb.Dataset(data = as.matrix(features_test), label = log(target_test))
params <- list(objective = "regression", metric = "rmse")
lightgbm_model <- lgb.train(params = params, data = dtrain,
                            nrounds = 100, verbose = -1)
lgbm_pred <- exp(predict(lightgbm_model, newdata = as.matrix(features_test)))
lgbm_mae <- MAE(lgbm_pred, testData$Price)
lgbm_rmse <- RMSE(lgbm_pred, testData$Price)
cat("LightGBM MAE:", lgbm_mae, "RMSE", lgbm_rmse)
```

# Conclusion
Let's see which model predicts the best.  
```{r}
rmse <- c(lr_rmse, svr_rmse, rf_rmse, xgb_rmse, lgbm_rmse)
mae <- c(lr_mae, svr_mae, rf_mae, xgb_mae, lgbm_mae)
performance <- round(rbind(rmse, mae),2)
colnames(performance) <- c("LR", "SVR", "RF", "XGB", "LightGBM")
data.frame(performance) %>% gt(rownames_to_stub = TRUE)
```
From the table above, Random Forest is the best model.

Next, we look at Feature importance of Random Forest, XGBoost and LightGBM.  
```{r}
# Random Forest
importance_scores <- importance(rf_model)
importance_df <- data.frame(Importance = importance_scores[, "%IncMSE"],
                            Feature = rownames(importance_scores))
importance_df <- importance_df[order(-importance_df$Importance), ]
top_10_features.rf <- importance_df[2:11,]$Feature

# XGBoost
importance_matrix <- xgb.importance(feature_names = colnames(features_train), model = xgb_model)

# Sorting the importance matrix by gain in descending order
top_features <- importance_matrix[order(-importance_matrix$Gain),]
# Selecting the top 10 features
top_10_features.xgb <- head(top_features, 10)$Feature

# LightGBM
feature_importance <- lgb.importance(model = lightgbm_model)
sorted_feature_importance <- feature_importance[order(-feature_importance$Gain), ]
top_10_features.lgb <- head(sorted_feature_importance, 10)$Feature

importance_comparison <- cbind(top_10_features.rf, top_10_features.xgb, 
                               top_10_features.lgb)
colnames(importance_comparison) <- c("RF", "XGB", "LightGBM")
rownames(importance_comparison) <- c(1:10)
data.frame(importance_comparison) %>% gt(rownames_to_stub = TRUE)
```
From table table above, we can conclude that the top 4 features to predict the price of a laptop are:  
1. Graphics_GB  
2. ppi  
3. Threads  
4. RAM_GB
