---
title: "<span style='color:	goldenrod;'>Laptop sales price prediction 2024</span>"
format: 
  html:
    toc: true
    theme: lux
    mainfont: "Times New Roman"
    code-fold: true
    fontsize: 1.3em
    linestretch: 1.7
title-block-banner: laptop_banner.jpg
editor: visual
editor_options:
  chunk_output_type: inline
execute: 
  echo: true
  warning: false
---

## Introduction to Dataset

This dataset contains information about various laptops, including specifications and prices, curated for sales price prediction. With features ranging from processor details to display specifications, this dataset serves as a valuable resource for analyzing trends and predicting laptop prices.

1.  Name: Name of the laptop model.

2.  Brand: Brand of the laptop.

3.  Price: Price of the laptop.

4.  Rating: Rating of the laptop.

5.  Processor_brand: Brand of the laptop's processor.

6.  Processor_name: Name of the laptop's processor.

7.  Processor_variant: Variant of the laptop's processor.

8.  Processor_gen: Generation of the laptop's processor.

9.  Core_per_processor: Number of cores per processor.

10. Total_processor: Total number of processors.

11. Execution_units: Number of execution units.

12. Low_Power_Cores: Number of low-power cores.

13. Energy_Efficient_Units: Indicates if the laptop has energy-efficient units.

14. Threads: Number of threads.

15. RAM_GB: RAM capacity of the laptop in gigabytes.

16. RAM_type: Type of RAM.

17. Storage_capacity_GB: Storage capacity of the laptop in gigabytes.

18. Storage_type: Type of storage.

19. Graphics_name: Name of the laptop's graphics.

20. Graphics_brand: Brand of the laptop's graphics.

21. Graphics_GB: Graphics capacity in gigabytes.

22. Graphics_integrated: Indicates if the laptop has integrated graphics.

23. Display_size_inches: Size of the laptop's display in inches.

24. Horizontal_pixel: Number of horizontal pixels.

25. Vertical_pixel: Number of vertical pixels.

26. ppi: Pixels per inch.

27. Touch_screen: Indicates if the laptop has a touch screen.

28. Operating_system: Operating system of the laptop.

```{r}
#| output: asis
library(Hmisc)
library(table1)
library(readr)
library(mice)
library(dplyr)
library(tidyselect)
library(DataExplorer)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(gt)
library(reticulate)
library(summarytools)


laptop <- read.csv("laptop.csv", na.strings = c("","NA"))
```

## Summary

```{r}
dfSummary(laptop)
```

## Missing Values

```{r}
#| output: asis
plot_missing(laptop, missing_only = TRUE)
```

## Missing Pattern

```{r fig.height=13, fig.width=13, message=FALSE, warning=FALSE}
mis_col <- colnames(laptop)[profile_missing(laptop)$num_missing > 0]
md.pattern_mine <- function (x, plot = TRUE, rotate.names = FALSE) 
{
    if (!(is.matrix(x) || is.data.frame(x))) {
        stop("Data should be a matrix or dataframe")
    }
    if (ncol(x) < 2) {
        stop("Data should have at least two columns")
    }
    R <- is.na(x)
    nmis <- colSums(R)
    R <- matrix(R[, order(nmis)], dim(x))
    pat <- apply(R, 1, function(x) paste(as.numeric(x), collapse = ""))
    sortR <- matrix(R[order(pat), ], dim(x))
    if (nrow(x) == 1) {
        mpat <- is.na(x)
    }
    else {
        mpat <- sortR[!duplicated(sortR), ]
    }
    if (all(!is.na(x))) {
        cat(" /\\     /\\\n{  `---'  }\n{  O   O  }\n==>  V <==")
        cat("  No need for mice. This data set is completely observed.\n")
        cat(" \\  \\|/  /\n  `-----'\n\n")
        mpat <- t(as.matrix(mpat, byrow = TRUE))
        rownames(mpat) <- table(pat)
    }
    else {
        if (is.null(dim(mpat))) {
            mpat <- t(as.matrix(mpat))
        }
        rownames(mpat) <- table(pat)
    }
    r <- cbind(abs(mpat - 1), rowSums(mpat))
    r <- rbind(r, c(nmis[order(nmis)], sum(nmis)))
    if (plot) {
        op <- par(mar = rep(0, 4))
        on.exit(par(op))
        plot.new()
        if (is.null(dim(sortR[!duplicated(sortR), ]))) {
            R <- t(as.matrix(r[1:nrow(r) - 1, 1:ncol(r) - 1]))
        }
        else {
            if (is.null(dim(R))) {
                R <- t(as.matrix(R))
            }
            R <- r[1:nrow(r) - 1, 1:ncol(r) - 1]
        }
        if (rotate.names) {
            adj <- c(0, 0.5)
            srt <- 90
            length_of_longest_colname <- max(nchar(colnames(r)))/2.6
            plot.window(xlim = c(-1, ncol(R) + 1), ylim = c(-1, 
                nrow(R) + length_of_longest_colname), asp = 1)
        }
        else {
            adj <- c(0.5, 0)
            srt <- 0
            plot.window(xlim = c(-1, ncol(R) + 1), ylim = c(-1, 
                nrow(R) + 1), asp = 1)
        }
        M <- cbind(c(row(R)), c(col(R))) - 1
        shade <- ifelse(R[nrow(R):1, ], mdc(1), mdc(2))
        rect(M[, 2], M[, 1], M[, 2] + 1, M[, 1] + 1, col = shade)
        for (i in 1:ncol(R)) {
            text(i - 0.5, nrow(R) + 0.3, colnames(r)[i], adj = adj, 
                srt = srt, cex = 0.6)
            text(i - 0.5, -0.8, nmis[order(nmis)][i], cex = 0.6, srt = srt)
        }
        for (i in 1:nrow(R)) {
            text(ncol(R) + 0.3, i - 0.5, r[(nrow(r) - 1):1, ncol(r)][i], 
                adj = 0, cex = 0.6)
            text(-0.3, i - 0.5, rownames(r)[(nrow(r) - 1):1][i], 
                adj = 1, cex = 0.6)
        }
        # text(ncol(R) + 1.3, -0.6, r[nrow(r), ncol(r)], cex = 0.6)
       # return(r)
    }

}
laptop[,mis_col] %>% md.pattern_mine(rotate.names = T)
```

# Exploratory Data Analysis (EDA)

We remove the 'X' column, as it solely contains the index, which we already have. Additionally, we also reconstruct the 'Price' column to 'Price_TWD', since the original values are in INR (Indian Rupees). The conversion to TWD (Taiwanese Dollars) should be done using the exchange rate of 0.39, as per the Google rate on 23/05/2024.

```{r}
laptop$Price_TWD <- laptop$Price * 0.39
laptop = subset(laptop, select = -c(Price, X))
```

Histogram of Price and log(Price)

```{r}
plot1 <- laptop %>%
  ggplot(aes(x = Price_TWD))+
  geom_histogram(binwidth = 10000, fill="lightblue",alpha=0.9)+
  ggtitle("Histogram of Price_TWD")

plot2 <- laptop %>%
  ggplot(aes(x = log(Price_TWD)))+
  geom_histogram(binwidth = 0.1, fill="lightblue",alpha=0.9)+
  ggtitle("Histogram of log(Price_TWD)")

grid.arrange(grobs = list(plot1, plot2), ncol = 2, nrow = 1)
```

Seeing from above result, we decided to use the log(Price) as our target y, as it fit the normal distribution better than the Price itself

```{r}
laptop$Price_TWD = log(laptop$Price_TWD)
```

To ensure data accuracy and consistency, we need to address and correct any typographical errors within the dataset. This step is crucial for maintaining the integrity of our analysis and ensuring that our insights are reliable.

```{r}
laptop[867, "RAM_type"] <- "LPDDR5X"
laptop[266, "Brand"] <- "Asus"
laptop$RAM_type <- gsub("LPDDRX4", "LPDDR4X", laptop$RAM_type)
# 
# laptop <- laptop %>%
#   filter(Processor_brand != 'Apple')

laptop[176, "RAM_type"] <- "LPDDR5"
laptop[296, "RAM_type"] <- "DDR4"
laptop[403, "RAM_type"] <- "DDR4"
laptop[552, "RAM_type"] <- "DDR4"
laptop[657, "RAM_type"] <- "DDR4"
laptop[664, "RAM_type"] <- "DDR4"
laptop[728, "RAM_type"] <- "DDR4"
laptop[774, "RAM_type"] <- "DDR4"
laptop[903, "RAM_type"] <- "LPDDR5"
laptop[999, "Graphics_brand"] = "NVIDIA"
laptop[laptop$Operating_system == "Android 11 OS", "Operating_system"] = "Prime OS"
laptop[laptop$Operating_system == "DOS 3.0 OS", "Operating_system"] = "DOS OS"
laptop[laptop$Operating_system == "Mac 10.15.3\t OS", "Operating_system"] = "Mac OS"
laptop[laptop$Operating_system == "Mac Catalina OS", "Operating_system"] = "Mac OS"
laptop[laptop$Processor_name == "Intel Core 3", "Processor_name"] = "Intel Core i3"
laptop[laptop$Processor_name == "Intel Core 5", "Processor_name"] = "Intel Core i5"
laptop[laptop$Processor_name == "Intel Core 7", "Processor_name"] = "Intel Core i7"
laptop[laptop$Processor_name == "Apple M2 Apple M2 Chip", "Processor_name"] = "Apple M2"
laptop[laptop$Processor_name == "Intel Core I3-1115G4", "Processor_name"] = "Intel Core i3"
laptop[laptop$Processor_name == "Apple M3 Max", "Processor_name"] = "Apple M3"
laptop[laptop$Processor_name == "AMD Athlon Pro", "Processor_name"] = "AMD Athlon"
laptop[laptop$Processor_name == "AMD Athlon Silver", "Processor_name"] = "AMD Athlon"
laptop[laptop$Processor_name == "eration Intel Core", "Processor_name"] = "Intel Core i5"
laptop[laptop$Processor_name == "Intel Pentium ", "Processor_name"] = "Intel Pentium"
laptop[laptop$Processor_name == "Intel Pentium Gold", "Processor_name"] = "Intel Pentium"
laptop[laptop$Processor_name == "Intel Pentium Silver", "Processor_name"] = "Intel Pentium"
laptop[laptop$Processor_name == "Intel  ", "Processor_name"] = "Intel Celeron"
laptop[laptop$Processor_name == "Intel Celeron Dual", "Processor_name"] = "Intel Celeron"
laptop[528, "Processor_name"] = "MediaTek Kompanio"
laptop[laptop$Processor_name == "Mediatek", "Processor_name"] = "MediaTek"
laptop[laptop$Processor_name == "HiSilicon  Kirin 9006C 9006C", "Processor_name"] = "Kirin 9006C"
laptop[laptop$Processor_name == "Microsoft SQ1 SQ1", "Processor_name"] = "Microsoft SQ1"
```

To enhance our understanding of the dataset and to perform a more effective exploratory data analysis (EDA), we will initially separate the dataset into numerical and categorical columns based on their data types. This classification will enable us to tailor our analysis techniques appropriately to each type of data.

## Numerical Columns

```{r}
numerical_columns <- laptop %>%
  select(where(is.numeric))

colnames(numerical_columns)
```

To understand the relationships between 'Price' and other numerical columns in the dataset, we will calculate the correlation coefficients. This analysis will help us identify how 'Price' interacts with other numerical variables, which can provide valuable insights into patterns and trends within the data.

```{r}
# Calculate correlation with Price_GBP
correlation <- sapply(numerical_columns, function(x) cor(x, laptop$Price_TWD, use = "complete.obs"))
# Sort the correlation values in descending order
sorted_correlation <- sort(correlation, decreasing = TRUE)
# Display the sorted correlation values excluding the first element (if needed)
sorted_correlation[-1]
```

**Correlation Coefficient Guidelines**

-   **0.70 to 1** : Strong positive correlation

-   **0.30 to 0.69** : Moderate positive correlation

-   **0.10 to 0.29** : Weak positive correlation

-   **0 to 0.09** : No correlation

```{r}
# Convert to data frame for ggplot2
correlation_df <- data.frame(
  Feature = names(correlation),
  Correlation = correlation
)

# Plotting with ggplot2
ggplot(correlation_df, aes(x = Correlation, y = reorder(Feature, Correlation))) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_vline(xintercept = c(0.1, 0.3, 0.7), color = c("red", "orange", "green"), linetype = "dashed") +
  theme_minimal() +
  labs(title = "Correlation between numerical columns and Price", x = "Correlation", y = "Feature")
```

Check the correlation across all numerical columns

```{r, fig.height=12,fig.width=12}
library(reshape2)
library(ggplot2)
# Compute the correlation matrix
correlation_matrix <- round(cor(numerical_columns, use = "pairwise.complete.obs"), 3)

# Melt the correlation matrix for ggplot2
melted_corr_matrix <- melt(correlation_matrix)

# Plotting with ggplot2
ggplot(data = melted_corr_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "black", linewidth = 0.5) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = value), color = "black", size = 4) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 7, hjust = 1)) +
  labs(title = "Correlation between numerical columns", x = "", y = "")
```

The observation that 'Energy_efficient_Units' and 'Low_Power_cores' have a correlation of 1 suggests that they may contain identical data. To confirm this, we will compare their unique values and examine their correlation further. If they are indeed the same, we will retain 'Energy_efficient_Units' and drop 'Low_Power_cores'.

Additionally, since 'Horizontal_pixel' and 'Vertical_pixel' are strongly correlated and can be inferred from 'ppi' (pixels per inch), we will simplify the dataset by dropping 'Horizontal_pixel' and 'Vertical_pixel'.

This approach will reduce redundancy and help streamline our analysis.

**Steps to implement:**

1.  **Compare 'Energy_efficient_Units' and 'Low_Power_cores':**

    -   Check unique values.

    -   Confirm correlation.

2.  **Drop redundant columns:**

    -   Remove 'Low_Power_cores'.

    -   Remove 'Horizontal_pixel'.

    -   Remove 'Vertical_pixel'.

```{r}
# Drop Low_Power_Cores column as it has the same information as Energy_Efficient_Units
laptop = subset(laptop, select = -c(Low_Power_Cores, Horizontal_pixel, Vertical_pixel))

# Remove Low_Power_Cores from numerical_columns
numerical_columns = subset(numerical_columns, select = -c(Low_Power_Cores))
```

## Categorical Columns

```{r}
categorical_columns <- laptop %>%
  select(where(is.character))

cat_cols <- colnames(categorical_columns)
cat_cols
```

```{r}
# Loop through categorical columns and print unique value counts
# Initialize an empty list to store the results
results <- list()

# Iterate through each categorical column
for (col in cat_cols) {
  # Calculate the number of unique values
  unique_values_count <- length(unique(laptop[[col]]))
  # Store the result in the list
  results[[col]] <- unique_values_count
}

# Convert the list to a dataframe
results_df <- data.frame(Column = names(results), UniqueValuesCount = unlist(results))

# Print the dataframe
results_df %>% gt()
```

Given the high cardinality of the features "Processor_name," "Processor_variant," and "Graphics_name," it is imperative to evaluate their contribution to the "Price." We will also ascertain if their influence is distinct from that of "Graphics_GB." For instance, consider using the Intel brand and Intel Core Ultra processor as examples for this assessment.

```{r}
# Processor_name vs. Graphics_GB
intel_df <- laptop %>% filter(Processor_brand == "Intel")
intel_df %>%
  ggplot(aes(x = Processor_name, y = Graphics_GB, fill = Processor_name))+
  geom_violin()+
  geom_jitter(color="black", size=0.4, alpha=0.9)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  ggtitle("Graphics_GB vs. Processor_name")+
  labs(x = "Processor_name")+
  theme(text = element_text(size = 10))

# Processor_name vs. RAM_GB
intel_df %>%
  ggplot(aes(x=Processor_name, y=RAM_GB, fill=Processor_name)) + 
  geom_violin() +
  geom_jitter(color="black", size=0.4, alpha=0.9)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  ggtitle("RAM_GB vs. Processor_name")+
  labs(x = "Processor_name")+
  theme(text = element_text(size = 10))

# Processor_name vs. Threads
intel_df %>%
  ggplot(aes(x=Processor_name, y=Threads, fill=Processor_name)) + 
  geom_violin() +
  geom_jitter(color="black", size=0.4, alpha=0.9)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  ggtitle("Threads vs. Processor_name")+
  labs(x = "Processor_name")+
  theme(text = element_text(size = 10))
```

From the plot, it is evident that these processor names also have the highest values for Graphics_GB, RAM_GB, and Threads. This indicates that these features, rather than the processor names themselves, are the primary drivers of the price.

```{r, fig.height=12, fig.width=10}
# Intel Core Ultra Processor variant vs RAM_GB
df_intel_core_ultra <- laptop %>% filter(Processor_name == "Intel Core Ultra")

# Plot 1: Intel Core Ultra Processor variant vs RAM_GB
p1 <- ggplot(df_intel_core_ultra, aes(x = Processor_variant, y = RAM_GB)) +
  geom_jitter() +
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10)) +
  ggtitle("Intel Core Ultra Processor variant vs RAM_GB")

# Plot 2: Intel Core Ultra Processor variant vs Price
p2 <- ggplot(df_intel_core_ultra, aes(x = Processor_variant, y = Price_TWD)) +
  geom_jitter() +
  geom_boxplot()+
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10)) +
  ggtitle("Intel Core Ultra Processor variant vs Price")

# Plot 3: Intel Core Graphics Name vs RAM_GB
p3 <- ggplot(df_intel_core_ultra, aes(x = Graphics_name, y = RAM_GB)) +
  geom_jitter() +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  ggtitle("Intel Core Graphics Name vs RAM_GB")

# Plot 4: Intel Core Graphics Name vs Price
p4 <- ggplot(df_intel_core_ultra, aes(x = Graphics_name, y = Price_TWD)) +
  geom_jitter() +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  ggtitle("Intel Core Graphics Name vs Price")

# Arrange the plots in a 2x2 grid with overall width and height
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

Since the plots for RAM_GB and Price_TWD are similar, we will remove the features "Processor_name," "Processor_variant," and "Graphics_name." The variance in these features is adequately captured by RAM_GB.

```{r}
laptop = subset(laptop, select = -c(Processor_name, Processor_variant, Graphics_name)) 
```

# Data Preprocessing

```{r}
# Re-isolate numerical and categorical columns
numerical_columns <- laptop %>%
  select(where(is.numeric)) %>%
  colnames()
categorical_columns <- laptop %>%
  select(where(is.character)) %>%
  colnames()
```

Since 'Name' contains information that can be explained by other features, we will drop it.

```{r}
X <- laptop

# Calculate correlation with Price_TWD
correlation <- sapply(numerical_columns, function(x) cor(X[[x]], X$Price_TWD, use = "complete.obs"))
# Sort the correlation values in descending order
sorted_correlation <- sort(correlation, decreasing = TRUE)

# Drop Name
X = subset(X, select = -c(Name, Processor_gen))
```

## Missing data imputation

We will be using MICE and method = 'pmm' to impute our missing data. And after we imputed the data, there are some rows that contains "space" value, which we will convert it to NA, and then delete the corresponding rows.

```{r}
#| warning: false
# MICE algorithm
temp_data <- mice(X, m=5, maxit=100, meth='pmm', seed=500, printFlag = FALSE)
X_imputed <- complete(temp_data, 1)

# Replace NA values in RAM_type based on Processor_brand being 'Apple'
X_imputed <- X_imputed %>%
  mutate(RAM_type = ifelse(is.na(RAM_type) & Processor_brand == 'Apple', 'Unified', RAM_type))

# Drop rows with any NA values
X_imputed[X_imputed == ""] = NA
X_imputed <- na.omit(X_imputed)

cat('Number of columns:', ncol(X_imputed), ', Number of rows:', nrow(X_imputed), '\n')
```

Now, let's check the summary again

```{r}
#| warning: false
dfSummary(X_imputed)
```

Since there is a duplicated row, let's just delete it.

```{r}
X_imputed = X_imputed %>% distinct()
dfSummary(X_imputed)
```

## Encoding

```{r}
categorical_columns <- X_imputed %>%
  select(where(is.character)) %>%
  colnames()
categorical_columns
```

Let's first delete some "Brand, Processor_brand, Graphics_brand" name that have frequency \<= 5, as having these values may be hard for us to split into training and testing data.

```{r}
# Use dplyr to count the frequency of each unique value in Column1
value_counts <- X_imputed %>%
  group_by(Brand) %>%
  summarise(count = n())%>%
  arrange(desc(count))

# Print the result
value_counts %>% gt()
```

```{r}
# Use dplyr to count the frequency of each unique value in Column1
value_counts <- X_imputed %>%
  group_by(Processor_brand) %>%
  summarise(count = n())%>%
  arrange(desc(count))

# Print the result
value_counts %>% gt()
```

```{r}
# Use dplyr to count the frequency of each unique value in Column1
value_counts <- X_imputed %>%
  group_by(Graphics_brand) %>%
  summarise(count = n())%>%
  arrange(desc(count))

# Print the result
value_counts %>% gt()
```

```{r}
# Use dplyr to count the frequency of each unique value in Column1
value_counts <- X_imputed %>%
  group_by(Operating_system) %>%
  summarise(count = n())%>%
  arrange(desc(count))

# Print the result
value_counts %>% gt()
```

```{r}
# Use dplyr to filter out rows where the frequency of values in Column1 is less than 2
X_filtered <- X_imputed %>%
  group_by(Brand) %>%
  filter(n() >= 5) %>%
  ungroup()  # To remove the grouping

X_filtered <- X_filtered %>%
  group_by(Processor_brand) %>%
  filter(n() >= 5) %>%
  ungroup()  # To remove the grouping

X_filtered <- X_filtered %>%
  group_by(Graphics_brand) %>%
  filter(n() >= 5) %>%
  ungroup()  # To remove the grouping

# X_filtered <- X_filtered %>%
#   group_by(Operating_system) %>%
#   filter(n() >= 3) %>%
#   ungroup()  # To remove the grouping

# # Print the filtered data frame
# print("Filtered data frame:")
head(X_filtered) %>% gt()
```

Based on the information in the Summary and the categorical columns shown above, we need to convert character data ('chr') into numerical format to prepare it for modeling. Let's examine the types of data in these categorical columns, particularly focusing on 'RAM_type' and 'Storage_type.' The other columns, such as those involving binary data or brand names, can be effectively transformed using binary encoding and one-hot encoding, respectively.

```{r, fig.height=12, fig.width=10}
p1 <- ggplot(X_filtered, aes(x = RAM_type, y = Graphics_GB)) +
  stat_summary(fun = mean, geom = "bar") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  ggtitle("RAM_type vs Graphics_GB")

p2 <- ggplot(X_filtered, aes(x = RAM_type, y = Price_TWD)) +
  stat_summary(fun = mean, geom = "bar") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10)) +
  ggtitle("RAM_type vs Price")

p3 <- ggplot(X_filtered, aes(x = Storage_type, y = Graphics_GB)) +
  stat_summary(fun = mean, geom = "bar") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10)) +
  ggtitle("Storage_type vs Graphics_GB")

p4 <- ggplot(X_filtered, aes(x = Storage_type, y = Price_TWD)) +
  stat_summary(fun = mean, geom = "bar") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10)) +
  ggtitle("Storage_type vs Price")

# Arrange the plots in a 2x2 grid with overall width and height
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

From above plots, we can see that DDR5 has the highest average Price compares to other RAM_type and SSD has highest average price compares to other Storage_type. But, as it don't really align with Price, we will categorize "RAM_type" and "Storage_type" as ordinal data, and rank them based on the time they are introduced.

**RAM_type** ranks (oldest to newest):

1.  DDR3, 2007

2.  LPDDR3, 2012

3.  DDR4, 2014

4.  LPDDR4, 2014

5.  LPDDR4X, 2016

6.  Unified, 2020

7.  DDR5, 2020

8.  LPDDR5, 2020

9.  LPDDR5X. 2021

10. DDR6, 2024

**Storage_type** ranks (oldest to newest):

1.  Hard Disk

2.  SSD

3.  Hard Disk + SSD

4.  NVMe SSD.

Types of Encoding we will use:

1.  Binary data: Binary Encoding for **Graphics_integreted** and **Touch_screen**

2.  Nominal data: One-hot Encoding for **Brand**, **Processor_brand**, **Graphics_brand** and **Operating_system**

3.  Ordinal data: Label Encoding for **RAM_type** and **Storage_type**

```{r}
# Convert to binary numeric values first
X_filtered$Graphics_integreted <- ifelse(X_filtered$Graphics_integreted == "True", 1, 0)
X_filtered$Touch_screen <- ifelse(X_filtered$Touch_screen == "True", 1, 0)

# Prepare factors for label encoding prior to one-hot encoding
# RAM_type with a specific order
level_order_ram <- c("DDR3", "LPDDR3", "DDR4", "LPDDR4", "LPDDR4X", "Unified", "DDR5", "LPDDR5", "LPDDR5X", "DDR6")
X_filtered$RAM_type <- factor(X_filtered$RAM_type, levels = level_order_ram)
X_filtered$RAM_type <- as.integer(X_filtered$RAM_type)

# Storage_type with a specific order
level_order_storage <- c(" Hard Disk", " SSD", "Hard Disk & SSD", " NVMe SSD")
X_filtered$Storage_type <- factor(X_filtered$Storage_type, levels = level_order_storage)
X_filtered$Storage_type <- as.integer(X_filtered$Storage_type)

# One-hot Encoding for the rest of the categorical variables
# Variables to encode
vars_to_encode <- c("Brand", "Processor_brand", "Graphics_brand", "Operating_system")
# Creating a model matrix for the selected variables
encoded_vars <- model.matrix(~ 0 + Processor_brand + Brand + Graphics_brand + Operating_system, data = X_filtered) # 
X_encoded <- cbind(X_filtered, encoded_vars)
# If original categorical variables should be dropped
X_encoded <- X_encoded[, !colnames(X_encoded) %in% vars_to_encode, drop = TRUE]

# Output the structure of the encoded data
cat('Number of columns:', ncol(X_encoded), ', Number of rows:', nrow(X_encoded), '\n')
```

Data Preprocessing is Done!! Let's check the Summary once again.

```{r}
#| warning: false
#| # Replace spaces with underscores in column names
names(X_encoded) <- gsub("\\s+", "_", names(X_encoded))
dfSummary(X_encoded)
write.csv(X_encoded, "laptop_encoded.csv", row.names=FALSE)
```

# Regression Models

We will do several regression models to predict the prices of laptops, but first we have to split the dataset into training and testing data.

```{r}
# Set the seed for reproducibility
set.seed(567)

# Randomly sample row indices
sample_indices <- sample(seq_len(nrow(X_encoded)), size = 0.8 * nrow(X_encoded))

# Split the data into training and testing sets
trainData <- X_encoded[sample_indices, ]
testData <- X_encoded[-sample_indices, ]

# Print the dimensions of the training and testing sets
cat("Training set dimensions:", dim(trainData), "\n")
cat("Testing set dimensions:", dim(testData), "\n")
```

We can generate the density plot of Price_TWD for both the training and testing datasets multiple times to verify if their distributions are similar.

```{r}
# Plot density

library(ggplot2)

ggplot() +
  geom_density(data = trainData, aes(x = Price_TWD, color = "Train Data", fill = "Train Data"), alpha = 0.328) +
  geom_density(data = testData, aes(x = Price_TWD, color = "Test Data", fill = "Test Data"), alpha = 0.328) +
  scale_color_manual(name = "Dataset", values = c("Train Data" = "red", "Test Data" = "green")) +
  scale_fill_manual(name = "Dataset", values = c("Train Data" = "blue", "Test Data" = "yellow")) +
  theme_minimal() +
  ggtitle("Density Plot of Train and Test Data Prices") +
  theme(legend.position = "right")
```

```{r}
# Set the seed for reproducibility
set.seed(123)

# Randomly sample row indices
sample_indices <- sample(seq_len(nrow(X_encoded)), size = 0.8 * nrow(X_encoded))

# Split the data into training and testing sets
trainData <- X_encoded[sample_indices, ]
testData <- X_encoded[-sample_indices, ]

write.csv(trainData, "trainData.csv", row.names=FALSE)
write.csv(testData, "testData.csv", row.names=FALSE)
# Print the dimensions of the training and testing sets
cat("Training set dimensions:", dim(trainData), "\n")
cat("Testing set dimensions:", dim(testData), "\n")
```

```{r}
# Plot density

library(ggplot2)

ggplot() +
  geom_density(data = trainData, aes(x = Price_TWD, color = "Train Data", fill = "Train Data"), alpha = 0.328) +
  geom_density(data = testData, aes(x = Price_TWD, color = "Test Data", fill = "Test Data"), alpha = 0.328) +
  scale_color_manual(name = "Dataset", values = c("Train Data" = "red", "Test Data" = "green")) +
  scale_fill_manual(name = "Dataset", values = c("Train Data" = "blue", "Test Data" = "yellow")) +
  theme_minimal() +
  ggtitle("Density Plot of Train and Test Data Prices") +
  theme(legend.position = "right")
```

## Linear Regression

```{r class.source = "fold-show"}
# Fit a linear model
model <- lm(Price_TWD ~ ., data = trainData)

# Print the model summary
summary(model)
```

Make predictions on the test data

```{r class.source = "fold-show"}
library(caret)
lr_predictions <- predict(model, newdata = testData)

# Calculate RMSE
lr_rmse <- sqrt(mean((exp(testData$Price_TWD) - exp(lr_predictions))^2))
cat("Root Mean Squared Error (RMSE):", lr_rmse, "\n")

# Calculate MAE
lr_mae <- mean(abs(exp(testData$Price_TWD) - exp(lr_predictions)))
cat("Mean Absolute Error (MAE):", lr_mae, "\n")
```

## SVM Regressor

```{r class.source = "fold-show"}
# Train the SVR model
library(e1071)
svr_model <- svm(Price_TWD ~ ., data = trainData, kernel = "linear")

svr_predictions <- predict(svr_model, newdata = testData)

# Calculate RMSE
svr_rmse <- sqrt(mean((exp(testData$Price_TWD) - exp(svr_predictions))^2))
cat("Root Mean Squared Error (RMSE):", svr_rmse, "\n")

# Calculate MAE
svr_mae <- mean(abs(exp(testData$Price_TWD) - exp(svr_predictions)))
cat("Mean Absolute Error (MAE):", svr_mae, "\n")
```

## Random Forest

```{r class.source = "fold-show"}
library(randomForest)
# Train the Random Forest model
rf_model <- randomForest(Price_TWD ~ ., data = trainData, importance = TRUE)

rf_predictions <- predict(rf_model, newdata = testData)

# Calculate RMSE
rf_rmse <- sqrt(mean((exp(testData$Price_TWD) - exp(rf_predictions))^2))
cat("Root Mean Squared Error (RMSE):", rf_rmse, "\n")

# Calculate MAE
rf_mae <- mean(abs(exp(testData$Price_TWD) - exp(rf_predictions)))
cat("Mean Absolute Error (MAE):", rf_mae, "\n")
```

## XGBoost

```{r class.source = "fold-show"}
library(xgboost)
# Creating feature matrix by excluding the target column explicitly
features_train <- trainData[, !colnames(trainData) %in% "Price_TWD"]
features_test <- testData[, !colnames(testData) %in% "Price_TWD"]

# Creating target vector
target_train <- trainData$Price_TWD
target_test <- testData$Price_TWD

# Train the XGBoost model
dtrain <- xgb.DMatrix(data = as.matrix(features_train), label = target_train)
dtest <- xgb.DMatrix(data = as.matrix(features_test), label = target_test)
xgb_model <- xgboost(data = dtrain, nrounds = 100, objective = "reg:squarederror", verbose = 0)

# Make predictions on the test data
xgb_predictions <- predict(xgb_model, dtest)

# Calculate RMSE
xgb_rmse <- sqrt(mean((exp(testData$Price_TWD) - exp(xgb_predictions))^2))
cat("Root Mean Squared Error (RMSE):", xgb_rmse, "\n")

# Calculate MAE
xgb_mae <- mean(abs(exp(testData$Price_TWD) - exp(xgb_predictions)))
cat("Mean Absolute Error (MAE):", xgb_mae, "\n")

# Assuming 'xgb_predictions' is your predicted data
prediction_data <- data.frame(Actual = testData$Price_TWD, Predicted = xgb_predictions)
prediction_data <- prediction_data[order(prediction_data$Actual),]

# Create a line plot
ggplot(prediction_data, aes(x = seq_along(Actual))) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +
  geom_line(aes(y = Predicted, color = "Predicted"), size = 1) +
  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) +
  labs(title = "Line Plot of Actual vs Predicted log(Prices)",
       x = "Index",
       y = "log(Price)") +
  theme_minimal()
```

## LightGBM

```{r class.source = "fold-show"}
#| warning: false
library(lightgbm)
# Creating feature matrix by excluding the target column explicitly
features_train <- trainData[, !colnames(trainData) %in% "Price_TWD"]
features_test <- testData[, !colnames(testData) %in% "Price_TWD"]

# Creating target vector
target_train <- trainData$Price_TWD
target_test <- testData$Price_TWD

# Create LightGBM datasets
dtrain <- lgb.Dataset(data = as.matrix(features_train), label = target_train)
dtest <- lgb.Dataset(data = as.matrix(features_test), label = target_test)
params <- list(
  objective = "regression",
  metric = "rmse"
)
lightgbm_model <- lgb.train(
  params = params,
  data = dtrain,
  nrounds = 100,
  verbose = -1
)

# Make predictions on the test data
lgbm_predictions <- predict(lightgbm_model, newdata = as.matrix(features_test))

# Calculate RMSE
lgbm_rmse <- sqrt(mean((exp(testData$Price_TWD) - exp(lgbm_predictions))^2))
cat("Root Mean Squared Error (RMSE):", lgbm_rmse, "\n")

# Calculate MAE
lgbm_mae <- mean(abs(exp(testData$Price_TWD) - exp(lgbm_predictions)))
cat("Mean Absolute Error (MAE):", lgbm_mae, "\n")
```

## Calculate Performance

```{r}
# Combine RMSE and mae values into vectors
rmse <- c(lr_rmse, svr_rmse, rf_rmse, xgb_rmse, lgbm_rmse)
mae <- c(lr_mae, svr_mae, rf_mae, xgb_mae, lgbm_mae)

# Create a matrix and set column names
performance <- rbind(rmse, mae)
colnames(performance) <- c("LR", "SVR", "RF", "XGB", "LightGBM")

# Convert the matrix to a data frame for easier manipulation
performance_df <- as.data.frame(performance)

# Format the values to two decimal places
performance_df[] <- lapply(performance_df, function(x) format(as.numeric(x), nsmall = 2, digits = 2, scientific = FALSE))

performance_df 
```

# Feature Importance

Plot feature importance from model we are using.

## Random Forest Feature Importance

```{r, fig.height=10, fig.width=7}
# Extract and prepare importance scores
importance_scores <- importance(rf_model)
importance_df <- data.frame(Importance = importance_scores[, "%IncMSE"], Feature = rownames(importance_scores))

# Sort and select the top 10 features
importance_df <- importance_df[order(-importance_df$Importance), ]
top_10_features <- head(importance_df, 10)

# Plot using ggplot2
library(ggplot2)
ggplot(top_10_features, aes(x = reorder(Feature, Importance), y = Importance, fill = Importance)) +
    geom_col() +
    coord_flip() +
    labs(title = "Random Forest Top 10 Important Features", x = "Features", y = "Importance") +
    theme_minimal()
```

## XGBoost Feature Importance

```{r, fig.height=10, fig.width=7}
# Extract feature importance based on gain
importance_matrix <- xgb.importance(feature_names = colnames(features_train), model = xgb_model)

# Sorting the importance matrix by gain in descending order
top_features <- importance_matrix[order(-importance_matrix$Gain),]
# Selecting the top 10 features
top_10_features <- head(top_features, 10)
# Print the top 10 features
# print(top_10_features)
# Plot the top 10 feature importances
xgb.plot.importance(importance_matrix = top_10_features, rel_to_first = TRUE, xlab = "Feature Importance")
```

## LightGBM Feature Importance

```{r}
feature_importance <- lgb.importance(model = lightgbm_model)
sorted_feature_importance <- feature_importance[order(-feature_importance$Gain), ]
# Plot feature importance using ggplot2
ggplot(sorted_feature_importance, aes(x = reorder(Feature, Gain), y = Gain, fill = Gain)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +  # Flips the axes for better readability
    labs(title = "Feature Importance from LightGBM Model", x = "Features", y = "Importance (Gain)") +
    theme_minimal()
```

# Python Session

## Catboost

```{python class.source = "fold-show"}
import pandas as pd
from catboost import CatBoostRegressor, Pool, EShapCalcType, EFeaturesSelectionAlgorithm

grid = {'iterations': [100, 150, 200],
        'learning_rate': [0.03, 0.1],
        'depth': [2, 4, 6, 8],
        'l2_leaf_reg': [0.2, 0.5, 1, 3, 5, 7, 9]}

catboost_model = CatBoostRegressor(random_state=4,
                         loss_function='RMSE',
                         eval_metric='RMSE', logging_level='Silent')
grid_search_result = catboost_model.grid_search(grid, r.features_train, r.target_train, verbose=20)
```

```{python}
import numpy as np
from sklearn.metrics import mean_squared_error, mean_absolute_error

catboost_best_params = grid_search_result['params']
catboost_best_model =  CatBoostRegressor(**catboost_best_params, random_state=4,
                         loss_function='RMSE',
                         eval_metric='RMSE',verbose=0)
                         
catboost_best_model.fit(r.features_train, r.target_train)
catboost_best_model.save_model('catboost_best_model.cbm') 
catboost_best_model_predict_arr = catboost_best_model.predict(r.features_test)
real_arr = r.target_test
catboost_rmse = np.sqrt(mean_squared_error(np.exp(catboost_best_model_predict_arr), np.exp(real_arr)))
catboost_mae = mean_absolute_error(np.exp(catboost_best_model_predict_arr), np.exp(real_arr))

print(f'RMSE Testing Loss: {catboost_rmse:.4f}')
print(f'MAE Testing Loss: {catboost_mae:.4f}')
```

## Catboost Feature Importance

```{python fig.cap="Catboost Feature Importance"}
import shap
import matplotlib.pyplot as plt

# Generate SHAP values using the TreeExplainer
explainer = shap.TreeExplainer(catboost_best_model)
shap_values = explainer(r.features_train)

# Create the beeswarm plot
fig = shap.plots.beeswarm(shap_values, show = False)

# Adjust layout to avoid text cut-off and make y-axis labels closer
plt.gca().margins(y=0.01)  # Adjust margins to make y-axis labels closer
plt.tight_layout(pad=1.0)  # Use tight_layout with a smaller pad

# Show the plot
plt.show()

```

## MLP

```{python}

import pandas as pd
import torch
import torch.nn as nn
import torch.optim as optim
from torch.utils.data import DataLoader, Dataset, TensorDataset
from sklearn.preprocessing import StandardScaler
# Assume 'data' is your pandas DataFrame containing the dataset

# Normalization
feature_names = r.features_train.columns
scaler = StandardScaler()
features_train = scaler.fit_transform(r.features_train)
features_train = pd.DataFrame(features_train, columns=feature_names)
features_test = scaler.transform(r.features_test)
features_test = pd.DataFrame(features_test, columns=feature_names)
target_train = np.exp(r.target_train)
target_test = np.exp(r.target_test)


# Define the groups
cpu_features = ['Core_per_processor', 'Total_processor',
       'Execution_units', 'Threads','Processor_brandAMD', 'Processor_brandApple', 'Processor_brandIntel', 'Energy_Efficient_Units']
gpu_features = ['Graphics_GB', 'Graphics_integreted','Graphics_brandApple','Graphics_brandIntel', 'Graphics_brandNVIDIA']
ram_features = ['RAM_GB', 'RAM_type']
other_features = ['BrandApple', 'BrandAsus', 'BrandAvita', 'BrandChuwi', 'BrandDell',
       'BrandGigabyte', 'BrandHonor', 'BrandHP', 'BrandInfinix', 'BrandLenovo',
       'BrandLG', 'BrandMicrosoft', 'BrandMSI', 'BrandSamsung', 'BrandUltimus',
       'BrandXiaomi', 'BrandZebronics','Operating_systemDOS_OS',
       'Operating_systemMac_OS', 'Operating_systemUbuntu_OS',
       'Operating_systemWindows_10_OS', 'Operating_systemWindows_11_OS', 'Rating']
storage_features = ['Storage_capacity_GB', 'Storage_type']
screen_features=['ppi', 'Touch_screen','Display_size_inches']
all_features = list(r.features_train.columns)
```

```{python}

# Custom Dataset class
class LaptopDataset(Dataset):
    def __init__(self, X, y):
        self.X = X
        self.y = y

    def __len__(self):
        return len(self.y)

    def __getitem__(self, idx):
        return {key: torch.tensor(val[idx], dtype=torch.float) for key, val in self.X.items()}, torch.tensor(self.y[idx], dtype=torch.float)

# Create DataLoaders
train_dataset = LaptopDataset({cat: features_train[cat].values for cat in features_train.columns}, target_train)
val_dataset = LaptopDataset({cat: features_test[cat].values for cat in features_test.columns}, target_test)

train_loader = DataLoader(train_dataset, batch_size=32, shuffle=True)
val_loader = DataLoader(val_dataset, batch_size=32, shuffle=False)

# Define the neural network with group-specific sub-networks
class LaptopPricePredictionModel(nn.Module):
    def __init__(self, num_features):
        super(LaptopPricePredictionModel, self).__init__()

        
        # Define sub-networks for each group
        self.cpu_net = self.create_sub_network(num_features['cpu'])
        self.gpu_net = self.create_sub_network(num_features['gpu'])
        self.screen_net = self.create_sub_network(num_features['screen'])
        self.ram_net = self.create_sub_network(num_features['ram'])
        self.storage_net = self.create_sub_network(num_features['storage'])
        self.other_net = self.create_sub_network(num_features['other'])
        self.all_net = self.create_sub_network(num_features['all'])

        # Combine all sub-networks' outputs
        combined_input_size = sum([net['output_size'] for net in [self.cpu_net, self.gpu_net, 
                                                                  self.screen_net, self.ram_net, 
                                                                  self.storage_net, self.other_net, self.all_net]])
        
        self.fc1 = nn.Linear(combined_input_size, 128)
        self.fc2 = nn.Linear(128, 64)
        self.fc3 = nn.Linear(64, 32)
        self.output = nn.Linear(32, 1)
        
        self.relu = nn.ReLU()
        self.gelu = nn.GELU()
        self.prelu = nn.PReLU()
    
    def create_sub_network(self, input_size):
        return {
            'net': nn.Sequential(
                nn.Linear(input_size, 64),
                nn.PReLU(),
                nn.Linear(64, 32),
                nn.PReLU()
            ),
            'output_size': 32
        }   

    def forward(self, x):
        # Process embeddings for categorical features
        
        # CPU sub-network
        cpu_inputs = torch.cat([x[feature].unsqueeze(1) for feature in cpu_features], 1)
        cpu_output = self.cpu_net['net'](cpu_inputs)
        
        # GPU sub-network
        gpu_inputs = torch.cat([x[feature].unsqueeze(1) for feature in gpu_features], 1)
        gpu_output = self.gpu_net['net'](gpu_inputs)
        
        # Screen sub-network
        screen_inputs = torch.cat([x[feature].unsqueeze(1) for feature in screen_features], 1)
        screen_output = self.screen_net['net'](screen_inputs)
        
        # RAM sub-network
        ram_inputs = torch.cat([x[feature].unsqueeze(1) for feature in ram_features], 1)
        ram_output = self.ram_net['net'](ram_inputs)
        
        # Storage sub-network
        storage_inputs = torch.cat([x[feature].unsqueeze(1) for feature in storage_features], 1)
        storage_output = self.storage_net['net'](storage_inputs)
        
        # Other sub-network
        other_inputs = torch.cat([x[feature].unsqueeze(1) for feature in other_features], 1)
        other_output = self.other_net['net'](other_inputs)
        
        all_inputs = torch.cat([x[feature].unsqueeze(1) for feature in all_features], 1)
        all_output = self.all_net['net'](all_inputs)
        
        # Concatenate all group-specific outputs
        combined_inputs = torch.cat([cpu_output, gpu_output, screen_output, ram_output, storage_output, other_output, all_output], 1)
        
        # Fully connected layers
        x = self.prelu(self.fc1(combined_inputs))
        x = self.prelu(self.fc2(x))
        x = self.prelu(self.fc3(x))
        x = self.output(x)
        
        return x


# Number of features in each group (adjust if your groups include more features)
num_features = {
    'cpu': len(cpu_features),
    'gpu': len(gpu_features),
    'screen': len(screen_features),
    'ram': len(ram_features),
    'storage': len(storage_features),
    'other': len(other_features),
    'all': len(all_features)
}

# Instantiate the model
mlp_model = LaptopPricePredictionModel(num_features)

# Define loss function and optimizer
criterion = nn.MSELoss()
optimizer = optim.Adam(mlp_model.parameters(), lr=0.001)

mlp_model
```

```{python}
# Training function
def train_model(model, train_loader, val_loader, criterion, optimizer, epochs=20):
    for epoch in range(epochs):
        model.train()
        train_loss = 0.0
        
        for inputs, target in train_loader:
            optimizer.zero_grad()
            outputs = model(inputs)
            loss = criterion(outputs, target.unsqueeze(1))
            loss.backward()
            optimizer.step()
            train_loss += loss.item() * len(target)
        
        train_loss /= len(train_loader.dataset)
        
        model.eval()
        val_loss = 0.0
        
        with torch.no_grad():
            for inputs, target in val_loader:
                outputs = model(inputs)
                loss = criterion(outputs, target.unsqueeze(1))
                val_loss += loss.item() * len(target)
        
        val_loss /= len(val_loader.dataset)
        if (epoch + 1) % 10 == 0 or epoch == epochs - 1:
            print(f'Epoch {epoch+1}/{epochs}, Training Loss: {train_loss:.4f}, Validation Loss: {val_loss:.4f}')

# Train the model
train_model(mlp_model, train_loader, val_loader, criterion, optimizer, epochs=80)

# Evaluate the model
mlp_model.eval()
with torch.no_grad():
    val_loss = 0.0
    for inputs, target in val_loader:
        outputs = mlp_model(inputs)
        loss = np.sqrt(criterion(outputs, target.unsqueeze(1)))
        val_loss += loss.item() * len(target)
    val_loss /= len(val_loader.dataset)

print(f'Model evaluation - Testing Loss: {val_loss:.4f}')
```

## Conclusion

```{r}
# Combine RMSE and mae values into vectors
rmse <- as.numeric(c(lr_rmse, svr_rmse, rf_rmse, xgb_rmse, lgbm_rmse, py$catboost_rmse, py$val_loss))

rmse <- round(rmse, digits = 3)
# Create a matrix and set column names
performance <- rmse
model <- c("LR", "SVR", "RF", "XGB", "LightGBM", "catboost", "nn")

# Convert the matrix to a data frame for easier manipulation
performance_df <- as.data.frame(cbind(model, performance))


performance_df %>% gt()
```
