---
title: "Group 6 - Final Report"
date: today
author: Tsung-Yu LEE
format:
 pdf:
    include-in-header:
      - text: |
         \usepackage{setspace,relsize}
mainfont: "Times New Roman"
toc: true
---

\newpage

To start the code, the following packages should be installed.

```{r setup, include = TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(Hmisc, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(table1, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(readr, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(mice, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(dplyr, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(tidyselect, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(DataExplorer, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(ggplot2, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(reshape2, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
library(VIM, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
```

# Introduction

## Data Source

The laptop dataset is from Kaggle [Laptop Sales Price Prediction](https://www.kaggle.com/datasets/siddiquifaiznaeem/laptop-sales-price-prediction-dataset-2024).

```{r}
laptop <- read.csv("C:/github_LTY/homework/Final Report/data/laptop.csv",
                   na.strings = c("","NA"))
```

## Summary Statistics

```{r}
#| output: asis
latex(describe(laptop), file = "", caption.placement = "top")
```

## Missing Values

```{r, fig.width=11}
plot_missing(laptop, missing_only = TRUE)
```
***Good***:


-   "Graphics_integrated", "Graphics_brand", "Graphics_name" have **2(0.2%)** missing values.


-   "Core_per_processor" has **12(1.18%)** missing values.


-   "RAM_type" has **22(2.16%)** missing values.


-   "Processor_variant" has **24(2.35%)** missing values. 


-   "Threads" has **48(4.71%)** missing values.


***OK***:


-   "Processor_gen" has **129(12.65%)** missing values.


-   "Execution_units" and "Total_processor" have **447(43.82%)** missing values.


***Bad***:


-   "Graphics_GB" has **652(63.92%)** missing values

## Data Visualization
```{r}
laptop[laptop$Brand == "Asus", "Brand"] = "ASUS"
sort(table(laptop$Brand))
```

```{r}

```


# Data Analysis

## Feature Engineering
```{r}
data <- laptop
```

## Split training and testing dataset
```{r}
set.seed(2024)
# smp.size = floor(0.8*nrow(data))
# train.index = sample(seq_len(nrow(data)), smp.size)
# train = data[train.index, ] # 80%
# test = data[-train.index, ] # 20%
```

## Model Fitting

### Linear Regression

### SVM Regressor
```{r}
library(e1071, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
# model <- svm(y ~ x , data)
# svr.pred = predict(model, data)
# mean(abs(data$y - svr.pred))
```

### Random Forest Regressor
```{r}
library(randomForest, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
```

### XGBoost

The parameters of XGBoost are listed below: ```max.depth```, ```eta```, ```nthread```, ```nrounds```, ```objective```: "binary:logistic", ```verbose```: Valid values are 0 (silent), 1 (warning), 2 (info), 3 (debug).
```{r}
library(xgboost, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
# xgb <- xgboost(data = train, max.depth = 2, eta = 1, nthread = 2, nrounds = 2, objective = "reg:absoluteerror", verbose = 0)
```

### LightGBM
```{r}
library(lightgbm, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
# lgbm <- lightgbm()
```

### Catboost
```{r}
library(catboost, warn.conflicts = FALSE, verbose = FALSE, quietly = TRUE)
```

# Final Results