---
title: "Laptop Price Prediction"
date: today
author: Tsung-Yu Lee
format:
 pdf:
    include-in-header:
      - text: |
         \usepackage{setspace,relsize}
mainfont: "Times New Roman"
toc: true
editor_options: 
  chunk_output_type: console
---

\newpage

# Introduction

## Data Source

The laptop dataset is from Kaggle [Laptop Sales Price Prediction](https://www.kaggle.com/datasets/siddiquifaiznaeem/laptop-sales-price-prediction-dataset-2024).

```{r, message=FALSE, warning=FALSE}
setwd("C:/github_LTY/homework/Final Report") # Set working directory.
laptop <- read.csv("data/laptop.csv", na.strings = c("","NA"))
```

Because "X" and "Name" are useless, so we delete these columns in advance.

```{r}
laptop <- subset(x = laptop, select = -c(X, Name))
```

## Summary Statistics

```{r, message=FALSE, warning=FALSE}
#| output: asis
library(Hmisc) # summary statistics
latex(describe(laptop), file = "", caption.placement = "top")
```

## Missing Values Proportion

```{r, fig.width=11, message=FALSE, warning=FALSE}
library(DataExplorer)
plot_missing(laptop, missing_only = TRUE)
```

From the figure above, we can discover that ```"Graphics_GB"``` has 63.92% missing values.

## Missing Pattern

Next, we want to realize the missing pattern of laptop data.

```{r, message=FALSE, warning=FALSE}
# Customize md.pattern function originally in mice package.
library(mice)
md.pattern.mine <- function(x, plot = TRUE, rotate.names = FALSE, cex = 1.0){
  if(!(is.matrix(x) || is.data.frame(x))){
    stop("Data should be a matrix or dataframe")
  }
  if(ncol(x) < 2){
    stop("Data should have at least two columns")
  }
  R <- is.na(x)
  # Calculate number of missing values in each variable.
  nmis <- colSums(R) 
  # create a matrix, and column is ordered by nmis.
  R <- matrix(R[, order(nmis)], dim(x))
  # Transform rows to strings consisting of 0(non-missing) and 1( missing).
  pat <- apply(R, 1, function(x) paste(as.numeric(x), collapse = ""))
  # Sort rows by their missing patterns.
  sortR <- matrix(R[order(pat), ], dim(x))
  
  if(nrow(x) == 1){mpat <- is.na(x)}
  # Keep non-duplicated rows.
  else{mpat <- sortR[!duplicated(sortR), ]}
  
  # obtain number of each missing pattern at LHS.
  rownames(mpat) <- table(pat)
  
  # abs(mpat - 1) obtain entries of matrix, 
  # (0, 1) means (missing, non-missing)
  # rowSums(mpat) obtain number of missing values in each row at RHS.
  r <- cbind(abs(mpat - 1), rowSums(mpat))
  # obtain number of missing values in each column at bottom center.
  # sum(nmis) sums values at bottom center.
  r <- rbind(r, c(nmis[order(nmis)], sum(nmis)))
  
  if (plot){
    op <- par(mar = rep(0, 4))
    # on.exit(par(op))
    plot.new()
    if(is.null(dim(sortR[!duplicated(sortR), ]))) {
      R <- t(as.matrix(r[1:nrow(r) - 1, 1:ncol(r) - 1]))
    }
    else{
      if (is.null(dim(R))) {
        R <- t(as.matrix(R))
      }
    R <- r[1:nrow(r) - 1, 1:ncol(r) - 1]
    }
    if(rotate.names){
      adj <- c(0, 0.5)
      srt <- 90
      length_of_longest_colname <- max(nchar(colnames(r)))/2.6
      plot.window(xlim = c(-1, ncol(R) + 1),
                ylim = c(-1, nrow(R) + length_of_longest_colname), asp = 1)
    }
    else{
      adj <- c(0.5, 0)
      srt <- 0
      plot.window(xlim = c(-1, ncol(R) + 1), 
                  ylim = c(-1, nrow(R) + 1), asp = 1)
    }
    M <- cbind(c(row(R)), c(col(R))) - 1
    shade <- ifelse(R[nrow(R):1, ], mdc(1), mdc(2))
    rect(M[, 2], M[, 1], M[, 2] + 1, M[, 1] + 1, col = shade)
    for(i in 1:ncol(R)){
      # Show colnames at the top center.
      text(i - 0.5, nrow(R) + 0.3, colnames(r)[i], adj = adj, 
          srt = srt, cex = cex)
      # Show number of missing values for each variable at the bottom center.
      text(i - 0.5, -0.8, nmis[order(nmis)][i], cex = cex, srt = srt)
    }
    for(i in 1:nrow(R)){
      # Show number of missing values for each subjects at the RHS.
      text(ncol(R) + 0.3, i - 0.5, r[(nrow(r) - 1):1, ncol(r)][i], 
          adj = 0, cex = cex)
      # Show number of subjects for each kind of missing pattern at the LHS.
      text(-0.3, i - 0.5, rownames(r)[(nrow(r) - 1):1][i], adj = 1, cex = cex)
    }
    # show total number of missing entries.
    # text(ncol(R) + 1.3, -0.6, r[nrow(r), ncol(r)], cex = cex) 
    # return(r)
  }
  else {
    return(r)
  }
}
```

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
library(dplyr)
mis_col <- colnames(laptop)[profile_missing(laptop)$num_missing > 0]
laptop[,mis_col] %>%
  md.pattern.mine(rotate.names = T, cex = 1.0)
```

From the missing pattern plot, we can observe that  
1. The missing values of "Total_processor" and "Execution_units" occur with each other.  

# Exploratory Data Analysis (EDA)

## Fix wrong values

First, we look at each categorical variables to check whether there are some typing errors.

```{r}
attach(laptop)
sort(table(Brand))
```

There are 157 subjects that are "Asus", and 1 subject that is "ASUS". We modify them to "ASUS".
```{r}
# replace "Asus" with "ASUS".
Brand <- gsub(pattern = "Asus", replacement = "ASUS", x = Brand)
```

Then, look at continuous variables.

```{r, fig.height=15, fig.width=15, message=FALSE, warning=FALSE}
library(reshape2)
library(ggplot2)
library(tidyverse)

numerical_columns <- laptop %>%
  select(where(is.numeric))
# Compute the correlation matrix using pairwise deletion.
cor_matrix <- round(cor(numerical_columns, use = "pairwise.complete.obs"), 3)

# Melt the correlation matrix for ggplot2
melted_cor_matrix <- melt(cor_matrix)

# Plotting with ggplot2
ggplot(data = melted_cor_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "black", linewidth = 0.5) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  geom_text(aes(label = value), color = "black", size = 6) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 10, hjust = 1)) +
  labs(title = "Correlation between numerical columns", x = "", y = "")
```

From the correlation heatmap, we can discover that  
1. ```"ppi"```, ```"Vertical_pixel"``` and ```"Horizontal_pixel"``` are highly correlated. We keep ```"ppi"```, and delete ```"Vertical_pixel"``` and ```"Horizontal_pixel"```.  
2. ```"Energy_Efficient_Units"``` and ```"Low_Power_Cores"``` are perfectly correlated ($\rho=1$). We keep ```"Energy_Efficient_Units"``` and delete ```"Low_Power_Cores"```.  
3. 
```{r}
# Delete "Graphics_GB" in the laptop data.
laptop <- subset(laptop, select = -c(Graphics_GB))
```

```{r}
MAE <- function(pred, true){
  return(mean(abs(pred-true)))
}
RMSE <- function(pred, true){
  return(sqrt(mean((pred-true)^2)))
}
```

